---
title: "Assignment 1"
author: "keerthi Tiyyagura"
date: "2023-09-24"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---
#Setting up the ISLR Package
#Giving the install.packages("ISLR") command in comments because it's averting me from knitting to pdf.
```{r}
library(ISLR)
```
#Calling the ISLR library
```{r}
summary(Carseats)
```
#Printing the summary of the Carseats dataset
```{r}
View(Carseats)
```
#It displays the data frame in the viewer to visually inspect the dataset(just used this command for my clarification)
```{r}
str(Carseats)
```
#Printing the Structure of the Carseats dataset
#This dataset contains 400 observations of 11 variables
```{r}
head(Carseats,8)
```
#Printing the first 8 rows of Carseats dataset
```{r}
max(Carseats$Advertising)
```
#Maximum value of Advertising attribute
```{r}
IQR(Carseats$Price)
```
#IQR(interquartile) of Price attribute
```{r}
plot(Carseats$Sales,Carseats$Price,main = "Running price during the sales",xlab = "Total Sales",ylab="Total Price")
abline(lm(Carseats$Price~Carseats$Sales),col="Red",lwd=2)
```
#Scatter Plot of Sales against Price
#This Plot shows the regression line of Sales vs Price and it has a negative slope ,implying a negative correlation ,and the value ranges from 0 to -1
#Also here we can observe that the regression line's quality of fit with points is good
```{r}
correlation<-cor.test(Carseats$Sales,Carseats$Price,method="pearson")
correlation
```
#Calculating the correlation of two attributes
#The correlation coefficient value is -0.444,it is indicating a negative direction.If one variable increases and other variable decreases the strength of the relationship will be moderate.

