---
title: "Assignment 3"
author: "Keerthi Tiyyagura"
date: "2023-10-29"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r}
# To create two variables X and Y
set.seed(2017)
X <- runif(100)*10
Y <- X*4+3.45
Y <- rnorm(100)*0.29*Y+Y
```
1.a)Plot Y against X. Include a screenshot of the plot in your submission. Using the File menu you can save the graph as a picture on your computer. Based on the plot do you think we can fit a linear model to explain Y based on X?
```{r}
# Plot Y against X
cor(X,Y)
```
```{r}
plot(X,Y)
# Yes,we can fit a linear model to explain Y based on X with a positive correlation.
```
b)Construct a simple linear model of Y based on X. Write the equation that explains Y based on X. What is the accuracy of this model?
```{r}
# Fit a linear model
model <- lm(Y~X)
summary(model)
```
```{r}
coefficients(model)
# Y=3.6108X+4.4655 is the model equation
# The accuracy of the above model is 65.17 percent.
```
c)How the Coefficient of Determination, R^2, of the model above is related to the correlation coefficient of X and Y?
```{r}
# Coefficient of determination
cor(X,Y)^2
# (Correlation Coefficient)ˆ2=Coefficient of Determination
```
2.
```{r}
head(mtcars)
```
2.a)James wants to buy a car. He and his friend, Chris, have different opinions about the Horse Power (hp) of cars. James think the weight of a car (wt) can be used to estimate the Horse Power of the car while Chris thinks the fuel consumption expressed in Mile Per Gallon (mpg),is a better estimator of the (hp). Who do you think is right? Construct simple linear models using mtcars data to answer the question.
```{r}
summary(lm(hp~wt,data = mtcars))
```

```{r}
summary(lm(hp~mpg,data = mtcars))
# By looking at the values, Chris is right, mpg had a high r square value of 60% compared to weight.
```

```{r}
# Fit simple linear models using mtcars dataset.
data(mtcars)
model.wt <- lm(hp~wt,data = mtcars)
model.mpg <- lm(hp~mpg,data = mtcars)
```

```{r}
summary(model.wt)
```

```{r}
summary(model.mpg)
```
2.b)Build a model that uses the number of cylinders (cyl) and the mile per gallon (mpg) values of a car to predict the car Horse Power (hp). Using this model, what is the estimated Horse Power of a car with 4 calendar and mpg of 22?
```{r}
summary(model <- lm(hp~cyl+mpg,data = mtcars))
```

```{r}
(model$coefficients[2]*4)+(model$coefficients[1])+(model$coefficients[3]*22)
```

```{r}
predict(model,data.frame(cyl=4,mpg=22),interval = "prediction",level = 0.85)
```

```{r}
# Installing package "mlbench"
library(mlbench)
data(BostonHousing)
```
3.a)Build a model to estimate the median value of owner-occupied homes (medv)based on the following variables: crime crate (crim), proportion of residential land zoned for lots over 25,000 sq.ft (zn), the local pupil-teacher ratio (ptratio) and weather the whether the tract bounds Chas River(chas). Is this an accurate model? (Hint check R 2 ) 
```{r}
model1 <- lm(medv~crim+zn+ptratio+chas,data = BostonHousing)
summary(model1)
# Due to the model's extremely low R square value of 36%, it is not particularly accurate.
```
3.b.1)Imagine two houses that are identical in all aspects but one bounds the Chas River and the other does not. Which one is more expensive and by how much?
```{r}
summary(model2 <- lm(medv~chas,data = BostonHousing))
```

```{r}
model2$coefficients
```

```{r}
(model2$coefficients[2]*0)+model2$coefficients[1]
```

```{r}
(model2$coefficients[2]*1)+model2$coefficients[1]
# The home with chas of 1 is more expensive than the house without chas of 0 with a value of 4.3 utility
```
3.b.2)Imagine two houses that are identical in all aspects but in the neighborhood of one of them the pupil-teacher ratio is 15 and in the other one is 18. Which one is more expensive and by how much?
```{r}
summary(model3 <- lm(medv~ptratio,data = BostonHousing))
```

```{r}
(model3$coefficients[2]*15)+model3$coefficients[1]
```

```{r}
(model3$coefficients[2]*18)+model3$coefficients[1]
#  The cost of the house with the ptratio of 15 is more expensive than the cost of the house with the ptratio 18 with a value of 6.47.
```
3.c)Which of the variables are statistically important (i.e. related to the house price)? Hint: use the p-values of the coefficients to answer.
```{r}
summary(model1)
#  If your p-value is low ( 0.05), you can reject the null hypothesis.
```
3.d)Use the anova analysis and determine the order of importance of these four variables.
```{r}
anova(model1)
# The order of importance of these four variables by comparing the p values are
# 1) crim
# 2) ptratio
# 3) zn
# 4) chas
```











